---
title: "Working with USGS Data and Functions"
author: "John Kemper & Matthew Ross"
date: "10/8/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
library(tidyverse)
library(dataRetrieval)
library(dygraphs) #Time series plotting
library(tidyr) 
library(xts) #Time series objects
library(ggthemes)


knitr::opts_chunk$set(echo = TRUE)
```

# Working with USGS data

## Download USGS data using dataRetrieval


### Non function oriented method

```{r,eval=F}
lincoln_q <- readNWISdv('06752260', parameterCd = '00060',
                        startDate = '1975-10-01',endDate = '2019-10-01') %>%
  rename(q_cfs = X_00060_00003,
         q_cd = X_00060_00003_cd) %>%
  mutate(site = 'lincoln')

#summary(lincoln_q)

elc_q <- readNWISdv('06752280', parameterCd = '00060',
                        startDate = '1975-10-01',endDate = '2019-10-01') %>%
  rename(q_cfs = X_00060_00003,
         q_cd = X_00060_00003_cd) %>%
  mutate(site = 'elc')


combo_data <- bind_rows(lincoln_q,elc_q)
summary(combo_data)
#summary(elc_q)
```


### Function oriented method


```{r}

q_downloader <- function(site_no = '06752260',
                         site_id = 'lincoln'){
  df <- readNWISdv(site_no, 
                   parameterCd = '00060',
                   startDate = '1950-10-01',
                   endDate = '2019-10-01') %>%
  rename(q_cfs = X_00060_00003,
         q_cd = X_00060_00003_cd) %>%
    mutate(site = site_id)
  return(df)
}


lincoln_q <- q_downloader(site_no = '06752260',
                          site_id = 'lincoln')
elc_q <- q_downloader(site_no = '06752280',
                      site_id = 'elc')
#
combo_data <- bind_rows(lincoln_q,elc_q)

summary(combo_data)
```


## Plot the discharge data

### Time series (xaxis = date)

```{r}

ggplot(combo_data,aes(x=Date,y=q_cfs,color=site)) + 
  geom_line() 


```

### Density plot 

```{r}
ggplot(combo_data,aes(x=q_cfs,color=site)) +
  geom_density() +
  scale_x_log10()

```


### Interactive graphs 

```{r,eval=F}
#Structore examination
#str(combo_data)
wide_data <- combo_data %>%
  select(-agency_cd,-site_no,-q_cd) %>%
  spread(key=site,value=q_cfs)

wide_xts <- xts(wide_data %>%
                  select(elc,lincoln),
                order.by = wide_data$Date)

dygraph(wide_xts)
```


### Function for interactive graph

```{r}

xts_maker <- function(df){
  wide_data <- df %>%
    select(-agency_cd,-site_no,-q_cd) %>%
    spread(key=site,value = q_cfs)

  wide_xts <- xts(wide_data %>%
                  select(-Date),
                order.by = wide_data$Date)
  
  return(wide_xts)
  
}


wide_xts <- xts_maker(combo_data)
wide_elc <- xts_maker(elc_q)

View(wide_data)

dygraph(wide_xts)
dygraph(wide_elc)
```


# In class work

## Write a function that makes ggplot time series graph of discharge

Here I want you to write your own function that adds custom 
themes to your normal ggplot call (so you don't have to write
out theme_few and theme(legend.position = ...)) over and over again. You may need to load the library `ggthemes`. 

This function call should generate a clean time series plot 
with a single command and a single option (which data frame to 
plot). Please plot the `combo_data` dataset here.


```{r}

###A function that plots discharge data
plot_builder <- function(df) {
  
  plot <- df %>%
    ggplot(.) +
    geom_line(aes(x = Date, y = q_cfs, color = site)) +
    theme_few() +
    scale_color_manual(values = c("black", "red")) +
    labs(y = "Discharge (m^3/s)", color = "Site") +
    theme(legend.position = c(0.1, 0.9))
    
  
  return(plot)
  
}


plot_builder(combo_data)


```


## Download discharge data

Here I want you to download data from any 2 other USGS site in the world using the `q_downloader` function we already wrote. If this 
function works, great! If not, you will need to investigate why? 
Even if your function call works, why is my function "brittle?"

Hint: hardcoding `q_cfs = X_00060_00003` could be dangerous

```{r}
###download discharge data for one site
q_DRKR <- q_downloader(site_no = "01589330",
                       site_id = "Dead Run Franklintown")

###plot the data
plot_builder(q_DRKR)


###download discharge for a second site
q_DR5 <- q_downloader(site_no = "01589312",
                      site_id = "Dead Run 5")

###plot the data for the second site
plot_builder(q_DR5)


```


## Joining your two datasets

Above, I combined datasets by stacking them on top of each other 
(bind_rows), but then we had to `spread` the data anyways because
`xts` and `dygraphs` prefer wide datasets. Here I want you 
to write a function that joins your two USGS datasets by calling one of the family of `dplyr::join`. 

Hint: be careful of having two `site` columns. 

```{r}


# dfspread <- spread(q_DRKR, key = site, value = q_cfs)

###Function to join data 
data_combo <- function(df1, df2) {
  
  q_both <- full_join(df1, df2, by = "Date", keep  = TRUE) %>%
    select(Date, site_no.x, site.x, q_cfs.x, site_no.y, site.y, q_cfs.y)
  
  
  
  
  return(q_both)
  
}


###call the function 
q_combo <- data_combo(q_DRKR, q_DR5)
View(q_combo)

```


## Plotting the data

My function above `xts_maker` merely preps the data for plotting 
a dygraph. Here, I want you to make a similar function that preps *and* plots a dygraph of your joined data. 


```{r}

##Function to prep data for a dynamic graph and make that dynamic graph
xts_doer <- function(df) {
  
  ###extract the names of the sites from the dataframe
  name1 <- sym(df$site.x[nrow(df)]) 
  name2 <- sym(df$site.y[nrow(df)]) 
  
  
  
  ###Select the relevant columns from the df we make in the join function above 
  ###and then rename the q column by site name
  q_plotboth <- df %>%
    select(Date, q_cfs.x, q_cfs.y) %>%
    rename(!!name1 := q_cfs.x, !!name2 := q_cfs.y)
  
    
  ###transform the dataframe into an xts dataframe
  q_xts <- xts(q_plotboth %>%
                 select(-Date),
               order.by = q_plotboth$Date)
  
  ####make the dynamic graph
  q_xts_graph <- dygraph(q_xts)
  
  return(q_xts_graph)
  
}

###Call the function to make the graph
xts_doer(q_combo)


```


